Model to study co-evolution of trade an culture

### 2015-03-27 article.tex
cultural change section

% 1. the second aspect is cultural change
% 2. an evolutionary framework has explored this
% 3. a large percentage of cultural products exchanged in an economic basis are neutral
% 4. explain neutrality, model of random copy.
% 5. from this perspective the interest of a product for an individual is directly related to its frequency
% 6. this can potentially affect the product as part of a market because offer/demand is related to frequency
% 7. at the same time, frequency will change depending on trade dynamics.

The second component of our model is focused on the dynamics of cultural change. They comprise the collection of processes that promote or inhibit the spread of information by social interaction within a population \cite[3]{boyd_origin_2005}. This evolutionary framework is being used by an increasing number of social scientists, fostering the development of transdisciplinary efforts designed to understand cultural evolution \cite{henrich_evolution_2003}. Archaeology has a rich tradition of evolutionary studies \cite{lycett_cultural_2015}, exploring concepts such as gene-culture coevolution \cite{burger_absence_2007}, diffusion processes \cite{fort_synthesis_2012}, phylogenetic analysis of material culture \cite{} or detection of rates of change in the archaeological record \cite{premo_cultural_2014}.

Of particular interest to our study are the scenarios where cultural traits are copied by unbiased transmission (i.e. traits which do not have a intrinsic positive or negative value for the individuals adopting them). This situation is often analysed following the genetic theory of neutral evolution \cite{neiman_stylistic_1995}, where the variation of traits is strictly explained by their frequency distribution in the population. There exist a large number of works using this model as the starting point to analyse patterns of cultural change in past societies \cite{lipo_neutralitystyle_2001,shennan_ceramic_2001,steele_ceramic_2010,kandler_nonequilibrium_2013,porcic_exploring_2014,crema_approximate_2014}.

Within this framework, the individual preferences for cultural traits (such as music taste or shapes of pottery) are copied from one individual to the other. Each individual is then adding some small variations to this cultural trait she just acquired. The individual will in turn continue to spread these cultural traits which will be further adopted by other individuals. Studies are then performed on the three key steps of the framework: the transmission, the variation and the differential replication. Interestingly the study of these elements is not only done thanks to humans but also thanks to other animals.

\xrc{I would use social learning instead of cultural transmission}

The transmission mechanisms are the means used to transfer cultural traits from one individuals to another. A large number of social learning schemes can be envisioned~\cite{heyes_social_1994} \xrc{a better reference would be \cite{henrich_evolution_2003}}. Notably, the mechanisms do not necessarily have to be as complex as teaching, since emulation and imitation can be sufficient to pass on certain cultural traits.

The variation mechanisms are at the heart of the creation of the creation of new cultural elements~\cite{obrien_variation_1990}. These may happen due to some errors during the copy process~\cite{schillinger_copying_2014}. In more recent culture we can witness the apparition of mutations with intentionality~\cite{ziman_technological_2003}\xrc{see also recombination \cite{sole_evolutionary_2013}}.

\xrc{Instead of differential replication I would use \emph{biased selection}}
The last element is the one that will interest us the most in this study: the differential replication between different cultural traits. If a cultural trait is found more often than the others in the population it means that something modified the rate at which it is adopted by each agent \xrc{beyond drift effects produced by random copy of neutral traits \cite{bentley_random_2004}}.
%The reproduction rate of ideas in a population can be characterise by different rates: directional, disruptive or stabilisation~\cite{rendell_cognitive_2011}. 
Multiple processes can be at play behind the reproduction \xrc{selection? transmission? reproduction is too biological} of cultural traits. Among them the three following will interest us due to their simplicity to test in an agent-based modelisation. First is the random copy of one trait from an agent to another agent. 
\todo{get a ref on that} \xrc{you can use \cite{bentley_random_2004}}
This type of mechanism reproduces well the changes in musical preferences \xrc{music here is odd; if an example is used it should be related to the topic of the paper}. Second are the the copy based on the frequency at which a cultural trait is observed. If a common trait is reproduced more often then we speak about ``conformity.'' If on the contrary the less common trait is preferred for reproduction we speak of ``rarity'' \xrc{nope, we speak of anti-conformism}. Finally a trait can be preferred depending on whom is carrying it. Among these preferences the prestige of the individual carrying it can constitute an explanation for the adoption foreign cultural traits in the Roman empire \xrc{this is known as model biased transmission}.

\xrc{Again, I should focus the lit.review on models of cultural change in at least ancient societies, this is too broad.}

>>>>

The second component of our model is focused on the dynamics of cultural change. They comprise the collection of processes that promote or inhibit the spread of information by social interaction within a population \cite[3]{boyd_origin_2005}. This evolutionary framework is being used by an increasing number of social scientists, fostering the development of transdisciplinary efforts designed to understand cultural evolution \cite{henrich_evolution_2003}. Archaeology has a rich tradition of evolutionary studies \cite{lycett_cultural_2015}, exploring concepts such as gene-culture coevolution \cite{burger_absence_2007}, diffusion processes \cite{fort_synthesis_2012}, phylogenetic analysis of material culture \cite{obrien_cladistics_2001} or detection of rates of change in the archaeological record \cite{premo_cultural_2014}.

A large percentage of these studies focus on the biases that affect the transmission of cultural traits, including the relevance of intrinsic traits and frequency dependance (e.g. conformism and anti-conformism). These analysis are often based in the genetic theory of neutral evolution \cite{neiman_stylistic_1995} and they take neutrality as a null-hypotheses (i.e. variation of traits is strictly explained by its frequency distribution in the population). There exist a large number of works using this model as the starting point to analyse patterns of cultural change in past societies \cite{lipo_neutralitystyle_2001,shennan_ceramic_2001,steele_ceramic_2010,kandler_nonequilibrium_2013,porcic_exploring_2014,crema_approximate_2014}.

Unbiased transmission generates a clear pattern of power law distributions that can be replicated with a simple random copying transmission mechanism \cite{bentley_random_2004}. Within this framework, an individual will copy the traits of a randomly chosen individual with a given probability. This copy can potentially introduce some errors in the acquired trait, which account for innovation processes. The individual will in turn continue to spread these cultural traits which will be further adopted by other individuals. This basic model can be enriched by several additional processes both in the innovation \cite{schillinger_copying_2014,sole_evolutionary_2013,ziman_technological_2003} and the transmission \cite{heyes_social_1994,henrich_evolution_2003}.

In particular, the trade of cultural products within a market system can potentially introduce a dynamics content-related bias linked to the price of the different products. This would also affect the frequency of the product within the population, which in turn would modify its price following a co-evolutionary dynamic.
